### 2018-11-26 tspl/instructions.tex
added instructions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Instructions for TSPL Exam
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt]{article}
\usepackage{a4,amssymb}
% \setlength{\oddsidemargin}{-1.5cm}
% \addtolength{\textwidth}{2cm}
% \addtolength{\textheight}{3cm}

\begin{document}
\pagestyle{empty}
\setcounter{page}{1}

\begin{center}
\large Types and Semantics for Programming Languages
\end{center}

\section*{Instructions}

\begin{enumerate}

\item  The exam lasts two hours.

\item Place your student identity card face-up on the desk in
  front of you.  The invigilator may come to check your identity, and
  in this case you may be asked to allow the invigilator to briefly
  use your computer.  The exam time has been calculated to allow time
  for such interruptions.

\item You may log into your computer as soon as you are ready to do so.

\item To download the exam paper, open a terminal window and 
type the following:
\begin{center}
   \texttt{getpapers}
\end{center}
This will create a subdirectory \texttt{tspl-pe} in your home directory,
containing the following files.
\begin{center}
\begin{tabular}{ll}
\texttt{tspl-pe/papers/exam.pdf}			& the exam \\
\texttt{tspl-pe/papers/instructions.pdf}		& these instructions \\
\texttt{tspl-pe/templates/Exam.lagda}			& exam template to edit \\
\texttt{tspl-pe/original\_templates/Exam.lagda}		& backup of exam template
\end{tabular}
\end{center}
The directories \texttt{tspl-pe/papers}, \texttt{tspl-pe/original\_templates} are
read only, but you may read and write \texttt{tspl-pe/templates}.

\item The textbook \emph{Programming Language Foundations in Agda} is available
by pointing your browser at:
\begin{center}
\texttt{file:///group/examreadonly/full/plfa/}
\end{center}
\item The Agda standard library is available by pointing your browser at:
\begin{center}
\texttt{file:///group/examreadonly/full/agda-stdlib/}
\end{center}
\item The Agda user manual is available by pointing your browser at:
\begin{center}
\texttt{file:///group/examreadonly/full/agda-docs/}
\end{center}
{\it (Note that internet access has been disabled.)}

\begin{center}
  \textbf{Do nothing further until the start of the exam is announced!}
\end{center}

\hfill \textit{Please Turn Over}
\newpage

\item When the start of the exam is announced, open the exam paper
\begin{center}
  \texttt{tspl-pe/papers/exam.pdf}
\end{center}
with the standard PDF viewer.

\item Change to the template directory
\begin{center}
  \texttt{cd tspl-pe/templates}
\end{center}
and edit the file
\begin{center}
  \texttt{Exam.lagda}
\end{center}
to include your answers, using Emacs or anything else suitable.
You are recommended to save your work on a regular basis.

\item Before submitting, make sure your file is processed by Agda
  correctly. Code which prevents the file from compiling should be
  made into comments. If you fail to solve part of a problem, you
  may get more credit if you indicate clearly which part you have
  not solved.

\item \emph{Please ensure before submission that the file}
  \texttt{Exam.lagda} \emph{contains your solutions to the exam.}  Submit
  your file using the command:
  \begin{center}
  \texttt{examsubmit Exam.lagda}
  \end{center}
  If you get an error, please check carefully that your file is called
  \texttt{Exam.lagda} and that you are in the same directory as this
  file. If you continue to have problems, please contact one of the
  invigilators.

Repeated submit commands are allowed, and will overwrite previous
submissions.  The last file submitted will be the one marked.

\item When the invigilators announce the end of the exam, you must
  submit and log out immediately.

\end{enumerate}

\end{document}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: